---
title: "Border Infrastructure Data"
author: "Gülzau, Fabian"
date: "`r format(Sys.time(), '%d/%m/%y')`"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here("text")) })
output: html_document
math: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkg_load, include = FALSE}
if(!require("xfun")) install.packages("xfun")
pkg_attach2("DiagrammeR", "grid", "tiff", "here")
```

### Introduction

This codebook provides a brief overview of the *Border Infrastructure Data*. The data
was collected as part of the research project [The Borders of the World: Processes
of De- and Rebordering in a Global Perspective](https://www.sfb1265.de/en/subprojects/the-borders-of-the-world-processes-of-de-and-rebordering-in-a-global-perspective/) (SFB 1265, C01, PI: Prof. Dr. Steffen Mau) which was funded
by the German Research Foundation, project number 290045248 - SFB 1265). For a 
more details please consult the following article: 

- Gülzau, F., & Mau, S. (2021). Walls, Barriers, Checkpoints, Landmarks, and "No-Man's-Land." 
A Quantitative Typology of Border Control Infrastructure. *Historical Social Research, 46*(3), 23-48. doi:10.12759/hsr.46.2021.3.23-48

The project developed a heuristic that measures the "physicality of borders" (Hassner 
& Wittenberg 2015: 162). We go beyond binary measures of border fortifications 
by classifying border architecture into five categories - from relatively to 
completely closed - that we describe respectively as "no-man's-land" borders, 
landmark borders, checkpoint borders, barrier borders, and fortified borders. See
Figure 1 below: 

```{r typology_fig, eval = TRUE, echo = FALSE, fig.cap = "*Figure 1. Typology of border infrastructures*"}
grid::grid.raster(tiff::readTIFF(here::here("figure", "Fig1 - Typology_logic.tif")))
```

### Approach

###### *from Gülzau & Mau (2021: 30-31)*

The data on border infrastructures were collected between April 2018 and October
2019. Our starting point was the “Direct Contiguity 3.2” data from the
Correlates of War Project that we used to identify all contiguous land borders
between nation states ($N_{country}=158$/$N_{border}=630$; Stinnett et al. 2002).[^1] 
For each country, we created a single document and gathered information on all the
land borders. In general, we relied on visual cues and additional evidence
from newspapers and digital sources. We looked for case studies in scientific
databases (e.g., Scopus) and conducted searches in digitized media archives.
We also examined satellite images of border crossing points using the Google
Maps API. In some cases, the image quality was too low, forcing us to use
alternative services such as Bing Maps. In addition to the visual information
included in satellite imagery, images of border posts were added where possible.
Lastly, we compared and enriched our coding with existing studies on
fortified borders (Avdan 2019; Hassner and Wittenberg 2015; Carter and Poast
2017; Jones 2012; Jellissen and Gottheil 2013).

```{r typology_diagram, eval = TRUE, echo = FALSE, fig.cap = "*Figure 2. Coding approach*"}
grViz(" 
digraph stakeholders {

graph [overlap = true, fontsize = 10, fontname = Montserrat]
rankdir = TB;

node [shape = box]
nodeA [label = 'Border\nbarrier'];
nodeB [label = 'Border\ncheckpoint'];
nodeC [label = 'At specific\npoints only?'];
nodeD [label = 'Dismantling by\nagreement'];

{
rank=same
node [shape = circle, fixedsize = TRUE, width = 1.3]
nodeE [label = 'No-man’s-land\nborder'];
nodeF [label = 'Landmark\nborder'];
nodeG [label = 'Checkpoint\nborder'];
nodeH [label = 'Barrier\nborder'];
nodeI [label = 'Fortified\nborder'];
}

nodeA->nodeB [label=' No', fontsize = 12];
nodeA->nodeC [label=' Yes', fontsize =12];
nodeB->nodeD [label=' No', fontsize = 12];
nodeB->nodeG [label=' Yes', fontsize =12];
nodeC->nodeH [label=' Yes', fontsize =12];
nodeC->nodeI [label=' No', fontsize = 12];
nodeD->nodeE [label=' No', fontsize = 12];
nodeD->nodeF [label=' Yes', fontsize =12];

}")
```

We followed a dyadic conceptualization of international borders by measuring
the infrastructure on each side of a mutual border line. Accordingly, a
shared border line that creates a “state couple” (Vallet and David 2014a) is
separated into two distinct observations. In this way, we account for borders
that are managed cooperatively as “bi-national institutions” (Longo 2017, 2)
and for borders that are places of conflict, which might be the case when incompatible
territorial designs meet. For instance, a state that is affected by a
fortified border could enter a race toward tighter borders by also installing
barriers or could de-escalate the situation by maintaining conventional
checkpoints. In addition, a dyadic approach enables us to include measurements
that capture the relationship between bordering countries such as differences
in the economic output or political system.

The typology enables us to map border infrastructures on a global scale,
addressing questions regarding the worldwide distribution of physical markers.
Nevertheless, our typology is not without limitations. First, borders and
their territorial designs have a history, but our typology only provides a crosssectional
view of current border infrastructures, as it was not possible to trace
the origin of each checkpoint. A case in point concerns African borders that
were drawn by colonial powers during the scramble for the continent and
securing colonial exploitation. Even today, the colonial past fuels border conflicts
and several border fences have been inherited from this past (Gülzau
and Mau 2021). Second, states maintain multiple border crossing points that
do not necessarily have the same material infrastructure along the whole border
line. Our measurement uses the highest level of border infrastructure at
a specific border line to characterize its entirety. For example, the Kenyan
government planned a border fence that was meant to cover the whole border
with Somalia. However, only one section at the border crossing point of
Mandera was eventually fenced (Galvin 2018). Accordingly, the border between
Kenya and Somalia is classified as a “barrier border,” although some
parts of the border are less protected. Lastly, our typology is limited to the
measurement of the physical infrastructure at a border line. However, border
infrastructures only regulate mobility when sufficiently monitored by personnel.
This is illustrated by former secretary of the U.S. Department of
Homeland Security, Janet Napolitano, who questioned the efficiency of border
walls saying “You show me a 50-foot wall and I’ll show you a 51-foot ladder
at the border” (Lacey 2011). 

##### Example: Turkmenistan - Uzbekistan

```{r dossier_fig, eval = TRUE, echo = FALSE, fig.cap = "*Figure 3. Example: Turkmenistan-Uzbekistan*", dpi = 300}
grid::grid.raster(tiff::readTIFF(here::here("figure", "Fig 3 - Example TKM-UZB.tif")))
```

[^1]: The COW dataset was adjusted by adding the border dyad between Nigeria and 
Chad, which became a land border due to the progressing aridification of Lake Chad.
In addition, two erroneous entries were corrected (United Arab Emirates-Qatar, and
Myanmar-Pakistan). Lastly, we excluded French overseas territories such as French 
Guyana. 