---
title: "Border Infrastructure Data"
author: "Gülzau, Fabian"
date: "`r format(Sys.time(), '%d/%m/%y')`"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = here::here("text")) })
output: html_document
math: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkg_load, include = FALSE}
if(!require("xfun")) install.packages("xfun")
pkg_attach2("tidyverse", "rio", "DiagrammeR", "grid", "tiff", "here", "kableExtra",
            "sjlabelled")
```

### Introduction

This codebook provides a brief overview of the *Border Infrastructure Data*. The data
was collected as part of the research project [The Borders of the World: Processes
of De- and Rebordering in a Global Perspective](https://www.sfb1265.de/en/subprojects/the-borders-of-the-world-processes-of-de-and-rebordering-in-a-global-perspective/) (SFB 1265, C01, PI: Prof. Dr. Steffen Mau) which was funded
by the German Research Foundation, project number 290045248 - SFB 1265). For more 
details on the dataset, please consult the following article: 

- [Gülzau, F., & Mau, S. (2021a). Walls, Barriers, Checkpoints, Landmarks, and "No-Man's-Land." 
A Quantitative Typology of Border Control Infrastructure. *Historical Social Research, 46*(3), 23-48. doi:10.12759/hsr.46.2021.3.23-48](https://www.sowi.hu-berlin.de/de/lehrbereiche/makro/News/hsr_46_2021_3_guelzau_mau_border-control-infrastructure.pdf)

The project developed a heuristic that measures the "physicality of borders" (Hassner 
& Wittenberg 2015: 162). We go beyond binary measures of border fortifications 
by classifying border architecture into five categories - from relatively to 
completely closed - that we describe respectively as "no-man's-land" borders, 
landmark borders, checkpoint borders, barrier borders, and fortified borders. 
Figure 1 provides a visual illustration of the categories.

```{r typology_fig, eval = TRUE, echo = FALSE, fig.cap = "*Figure 1. Typology of border infrastructures*"}
grid.raster(readTIFF(here("figure", "Figure 1 - Typology of border infrastructures.tif")))
```

### Approach

###### (*from Gülzau & Mau 2021a: 30-31*)

The data on border infrastructures were collected between April 2018 and October
2019. Our starting point was the “Direct Contiguity 3.2” data from the
Correlates of War Project that we used to identify all contiguous land borders
between nation states ($N_{country}=158$, $N_{border}=630$; Stinnett et al. 2002).[^1] 
For each country, we created a single document and gathered information on all the
land borders. In general, we relied on visual cues and additional evidence
from newspapers and digital sources. We looked for case studies in scientific
databases (e.g., Scopus) and conducted searches in digitized media archives.
We also examined satellite images of border crossing points using the Google
Maps API. In some cases, the image quality was too low, forcing us to use
alternative services such as Bing Maps. In addition to the visual information
included in satellite imagery, images of border posts were added where possible.
Lastly, we compared and enriched our coding with existing studies on
fortified borders (Avdan 2019; Carter and Poast 2017; Hassner and Wittenberg 2015; 
Jellissen and Gottheil 2013; Jones 2012; Linnell et al. 2016).[^2] Figure 2 
illustrates our coding approach. 

```{r typology_diagram, eval = TRUE, echo = FALSE, fig.cap = "*Figure 2. Coding approach*"}
grViz(" 
digraph stakeholders {

graph [overlap = true, fontsize = 10, fontname = Montserrat]
rankdir = TB;

node [shape = box]
nodeA [label = 'Border\nbarrier'];
nodeB [label = 'Border\ncheckpoint'];
nodeC [label = 'At specific\npoints only?'];
nodeD [label = 'Dismantling by\nagreement'];

{
rank=same
node [shape = circle, fixedsize = TRUE, width = 1.3]
nodeE [label = 'No-man’s-land\nborder'];
nodeF [label = 'Landmark\nborder'];
nodeG [label = 'Checkpoint\nborder'];
nodeH [label = 'Barrier\nborder'];
nodeI [label = 'Fortified\nborder'];
}

nodeA->nodeB [label=' No', fontsize = 12];
nodeA->nodeC [label=' Yes', fontsize =12];
nodeB->nodeD [label=' No', fontsize = 12];
nodeB->nodeG [label=' Yes', fontsize =12];
nodeC->nodeH [label=' Yes', fontsize =12];
nodeC->nodeI [label=' No', fontsize = 12];
nodeD->nodeE [label=' No', fontsize = 12];
nodeD->nodeF [label=' Yes', fontsize =12];

}")
```

In particular, we distinguish borders that are fortified (e.g. fences, walls, 
barbed-wire, and landmines) from more open border lines. The fortified borders 
are then coded according to the extent of border fortifications. States with 
*fortified borders* install obstacles that are meant to prevent unauthorized 
mobility along the total length of a border line, while states with *barrier borders*
use physical barriers along specific locations that make it difficult to avoid 
inspections. The remaining categories differ by the use of designated checkpoints. 
*Checkpoint borders* are characterized by border posts at major border crossing 
points with the purpose of stopping and controlling travelers. In contrast, states
that have landmark borders and no-man's-land borders are not relying border posts
to monitor cross-border mobility. However, in the case of *landmark borders*, 
states have agreed to abolish regular controls in order to boost the cross-border
flows of goods and people, while *no-man's-land borders* are often found in remote 
regions such as deserts or jungles, which are difficult to access by state agents.

We followed a dyadic conceptualization of international borders by measuring
the infrastructure on each side of a mutual border line. Accordingly, a shared 
border line that creates a “state couple” (Vallet and David 2014) is 
separated into two distinct observations. In this way, we account for borders
that are managed cooperatively as “bi-national institutions” (Longo 2017, 2)
and for borders that are places of conflict, which might be the case when incompatible
territorial designs meet. For instance, a state that is affected by a
fortified border could enter a race toward tighter borders by also installing
barriers or could de-escalate the situation by maintaining conventional
checkpoints. In addition, a dyadic approach enables us to include measurements
that capture the relationship between bordering countries such as differences
in the economic output or political system.

Using the case of the border between Turkmenistan and Uzbekistan, figure 3 gives
an example of information that was used to classify borders. The respective border
is coded as a *fortified border* because existing data sets and policy reports 
indicate that large stretches of the border are equipped with fences. Using 
satellite images, the existence of a border fence was confirmed. 

##### Example: Turkmenistan - Uzbekistan

```{r dossier_fig, eval = TRUE, echo = FALSE, fig.cap = "*Figure 3. Example: Turkmenistan-Uzbekistan*", dpi = 300}
grid.raster(readTIFF(here("figure", "Fig 3 - Example TKM-UZB.tif")))
```

The typology enables us to map border infrastructures on a global scale,
addressing questions regarding the worldwide distribution of physical markers.
Nevertheless, our typology is not without limitations. First, borders and
their territorial designs have a history, but our typology only provides a crosssectional
view of current border infrastructures, as it was not possible to trace
the origin of each checkpoint. A case in point concerns African borders that
were drawn by colonial powers during the scramble for the continent and
securing colonial exploitation. Even today, the colonial past fuels border conflicts
and several border fences have been inherited from this past (Gülzau
and Mau 2021b). Second, states maintain multiple border crossing points that
do not necessarily have the same material infrastructure along the whole border
line. Our measurement uses the highest level of border infrastructure at
a specific border line to characterize its entirety. For example, the Kenyan
government planned a border fence that was meant to cover the whole border
with Somalia. However, only one section at the border crossing point of
Mandera was eventually fenced (Galvin 2018). Accordingly, the border between
Kenya and Somalia is classified as a “barrier border,” although some
parts of the border are less protected. Lastly, our typology is limited to the
measurement of the physical infrastructure at a border line. However, border
infrastructures only regulate mobility when sufficiently monitored by personnel.
This is illustrated by former secretary of the U.S. Department of
Homeland Security, Janet Napolitano, who questioned the efficiency of border
walls saying “You show me a 50-foot wall and I’ll show you a 51-foot ladder
at the border” (Lacey 2011). 

### Variables

```{r load_data, eval = TRUE, echo = FALSE}
border.df <- import(here("data", "border-infrastructure-data.rds"))

tibble(
  Variables = colnames(border.df),
  Description = get_label(border.df),
  Labels = get_labels(border.df, values = "p")
  ) %>%
  mutate(Labels = ifelse(lengths(Labels) > 5, "", Labels)) %>%
  kbl() %>%
  kable_styling()
```

### Bibliography

Avdan, N. (2019). *Visas and Walls. Border Security in the Age of Terrorism*. Philadelphia: University of Pennsylvania Press.

Carter, D. B., & Poast, P. (2017). Why Do States Build Walls? Political Economy, Security, and Border Stability. *Journal of Conflict Resolution, 61*(2), 239-270. 

Galvin, M. (2018). Combating Terrorism on the Somalian Border: The Improbable Kenyan Dream? *Global Challenges*(4). 

Gülzau, F., & Mau, S. (2021a). Walls, Barriers, Checkpoints, Landmarks, and "No-Man's-Land." A Quantitative Typology of Border Control Infrastructure. *Historical Social Research, 46*(3), 23-48. doi:10.12759/hsr.46.2021.3.23-48

Gülzau, F., & Mau, S. (2021b). Walls, Barriers, Checkpoints, No-man's-land. A Typology of Border Infrastructure on the African Continent. *Comparativ. Zeitschrift für Globalgeschichte und vergleichende Gesellschaftsforschung, 31*(3/4), 411-438. doi:10.26014/j.comp.2021.03-04.07

Hassner, R. E., & Wittenberg, J. (2015). Barriers to Entry. Who Builds Fortified Boundaries and Why? *International Security, 40*(1), 157-190. doi:10.1162/ISEC_a_00206

Jellissen, S. M., & Gottheil, F. M. (2013). On the utility of security fences along international borders. *Defense & Security Analysis, 29*(4), 266-279. doi:10.1080/14751798.2013.842707

Jones, R. (2012). *Border Walls: Security and the War on Terror in the United States, India, and Israel*. London: Zed Books.

Lacey, M. (2011, July 19, 2011). Arizona Officials, Fed Up With U.S. Efforts, Seek Donations to Build Border Fence. *The New York Times*. Retrieved from https://www.nytimes.com/2011/07/20/us/20border.html

Linnell, J. D. C., Trouwborst, A., Boitani, L., Kaczensky, P., Huber, D., Reljic, S., . . . Breitenmoser, U. (2016). Border Security Fencing and Wildlife: The End of the Transboundary Paradigm in Eurasia? *PLOS Biology, 14*(6), e1002483. doi:10.1371/journal.pbio.1002483

Longo, M. (2017). *The Politics of Borders: Sovereignty, Security, and the Citizen after 9/11*. Cambridge: Cambridge University Press.

Stinnett, D. M., Tir, J., Diehl, P. F., Schafer, P., & Gochman, C. (2002). The Correlates of War (Cow) Project Direct Contiguity Data, Version 3.0. *Conflict Management and Peace Science, 19*(2), 59-67. doi:10.1177/073889420201900203

Vallet, E., & David, C.-P. (2014). Good Fences for Good Neighbours? States, Couples and Walls. In B. Vassort-Rousset (Ed.), *Building Sustainable Couples in International Relations: A Strategy Towards Peaceful Cooperation* (pp. 66-87). London: Palgrave Macmillan UK.

[^1]: The COW dataset was adjusted by adding the border dyad between Nigeria and 
Chad, which became a land border due to the progressing aridification of Lake Chad.
In addition, two erroneous entries were corrected (United Arab Emirates-Qatar, and
Myanmar-Pakistan). Lastly, we excluded French overseas territories such as French 
Guyana. 

[^2]: The image database is available upon request. 